\TheTitle

### 2017-07-14 paper/study.tex
Editing first half of study

The purpose of the study is to quantitatively and qualitatively evaluate
the effectiveness of \tool compared to the DAG-based visualizations
found in Gitk or the git command line. The study is broken into three
task-based sections; conceptual tasks, summarization tasks, and
participant opinions. The study was performed during June of 2017
%in Kingston ON,
in a controlled environment running Ubuntu 14.04.
Participants were allowed to use both Gitk and the git command line
tools together, which we will refer to as the git tools or as Gitk. Then
they were asked to perform the same tasks using \tool.

\dmg{I thought you had randomized the order of tool use? also clarify that they were done with different commits to
  avoid bias}

% \evan{Do I need to say where the study took place? I see it showing up
%   in other papers, but it makes it kind of obvious about the lab I'm
%   talking about, doesn't it?}

The primary goal of the study is to determine if \tool is more capable
of providing the participants of our study with a better conceptual
understanding of how a commit is integrated into the kernel, and a
better summarization of various metrics about the files and authors
involved with the integration. The conceptual questions are to determine
if the DAG-based visualizations (gitk and git command line) provide enough information to determine how a
commit, and any related commits, are merged into a project. \tool
displays this information directly in the form of a tree, so we will not
be performing a comparison on this section, only evaluating the results
from Gitk. The summarization tasks compare \tool and Gitk, with the
participants using one tool followed by the other on two commits from
merge-trees of differeing sizes. In both of these sections, the
participant is given a task that they are to find an answer to; we
evaluate the time it took, if the answer was correct, and how far from
the right answer they were. The user opinions provide us with a clear
comparison between the tools from the point-of-view of the participant,
providing us information about which aspects of each tool they prefered.

@@

% \dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}

\dmg{I don't like how you talk about the DAG comprehension. Few paragraphs above I replaced it
  with the DAG-based tools, which might be a better description, since you are evaluating tool
vs tool}

@@

\textbf{The conceptual tasks} are designed to gain better insight on the
comprehension of the DAG, and determine if the DAG is sufficient for
understanding what other commits are related to a given commit, and how
those commits are merged. The tasks are outlined in
Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
using the git tools. T1 has the participant draw a diagram of the merge
tree, which will be used to answer the other two questions. This enables
us to visually see issues with comprehending the DAG visualizations.
Participants are given 10 minutes to perform this task, which also is
used to familiarize the participant with the git tools interface, if
they are not already, and with the set of commits they will be working
with. T2 and T3 are drawn directly from the diagram from T1, giving
quantitative metrics to measure the comprehension. T2 asks the
participant to verify the number of commits, and T3 asks for the number
of merges. The three tasks are performed on both commits before
continuing to the summarization tasks.

\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
are designed to compare the ability of the participants to summarize
information about the two commits used in the study when using git tools\dmg{be consistent, either you say dag-based
  tools, or git tools, either one is fine, but be consistent}
versus \tool. The tasks are broken into four sets based on what is being
summarized: merges, authorship, files, and modules. The order that the
participants perform each task is randomized within each task set, and
the order of the task sets are then randomized with the exception of the
merges task set which will always come first. We measure and evaluate
three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

\evan{Is this better, or do you still want me to move the information
  about the randomization to the bottom?}
\dmg{I didn't see anything about choosing what commits to inspect here... but this section looks good}

While the merge task set is part of the summarization task group, it
works less with the summarization of the commits and more with
comprehension of the models. The task is designed with two goals in
mind; first we want to measure and define which commits and which merges
the participant has defined to be in the merge tree; second to help
solidify which commits and merges will be used in the merge tree.

T4 provides a concrete series of merges that the participant believes to
be integrating the commit into the repository. T5 provides a concrete
set of commits that are related to the commit. We place these tasks
first as the answers to these will be necessary for the participant to
answer the tasks that follow, specifically finding the merge into the
master branch. T6 and T7 are related to the authorship at merges,
determining how many authors are involved and who made the most changes.
T8 and T9 are related to the files that were modified in the merge,
determining how many files were modified and which file had the most
changes. T10 is related to the modules that were modified in the merge.
Modules are not explicitly defined in git, but are a property of the
Linux repository that we noticed while looking at various commits. To
find the modules, take the text up to the first colon in the git log
preview. For the log preview \textit{``ALSA: kernel docs: fix
  sound/core/ kernel-doc''} the module is \textit{``ALSA''}, e.g.

The accuracy metrics are measured slightly differently between tasks due
to different requirements and structures in the answer. In task T4, both
the order and the merges are important. We define the distance to be the
edit distance with insertion, deletion, and transposition at unit cost.
The accuracy is the minimum edit distance necessary to make the answer
correct. T5, T9, and T10 are concerned with the set of elements in the
response; the order is not important. We use the Levenshtein distance,
with insertion, deletion, and substitution operations of unit cost, from
the provided answer to the correct answer. The answers to T6 and T8 are
single number values; the distance is measured as the absolute value of
the difference between the provided answer and the correct answer. T7
has only a single correct answer, thus the distance would either be zero
or one. For this reason, it is acceptable to omit the accuracy metric
from this task. The correctness metric is simply whether the answer was
correct or not; an accuracy of zero indicates a correct answer.

@@

We include two commits in the study to detect differences between merge
tree sizes. The order that the commits are presented to each participant
is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
nodes non-root nodes. 75\% of the trees contain at least 51 nodes, and
the largest tree contains 7217 nodes.

From this information, we selected one tree with a single commit, and
one tree with seven other nodes to represent a small tree and a
medium-sized tree. A majority of the trees are flat, where all nodes
merge directly into the root node. Of the 8031 trees containing at least
seven nodes, only 593 contained at least a single internal merge node.
Trivially, trees with a single node cannot have any internal nodes, and
of the 624 trees with seven nodes, only 135 contained at least one inner
node. To increase the complexity of the medium-sized tree, we randomly
selected one of the 135 trees to represent the medium tree.

From the 2008 trees containing only a single node, we selected one at
random, and trivially selected the only commit in the tree. From the
trees containing seven nodes, we selected one randomly, then randomly
selected one of the commits in the tree. We selected commit
\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
a single node, which we will refer to as \comA, and commit
\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
seven nodes, which we will refer to as \comB. A visual representation
for these trees is shown in Figure~\ref{fig:study_commits}.

\begin{figure}[bpt]
  \centering
  \begin{tabular}{ m{1.5cm} m{3cm} }
    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
  \end{tabular}
  \caption{The merge trees used in the user study,
    containing one and seven commits, respectively.}
  \label{fig:study_commits}
\end{figure}

@@

The study was conducted with 12 participants, all of whom were
masters, PhD, or post-doc researchers in the field of software
engineering. The participants had between 6 months and 10 years
experience with git, with the median being 3.5 years. Most of the
participants had additional experience with SVN and CVS.\@ One of the
participants in the study worked as a release engineer, studying merge
practices to determine the best way to merge branches while minimizing
the number of merge conflicts, on SVN repositories. The participants
worked with repositories ranging from around 10 commits up to 38000
commits, with the median being 350 commits. Two of the participants had
never collaborated with anyone in a repository, while the rest had some
experience with repositories being modified by multiple people, with the
most being 219. The median number of collaborators was four.

Five participants had only worked with personal and academic
repositories. Two participants had worked with school and
professional repositories, and two had only worked with professional
repositories. One participant worked with personal, academic, and
professional repositories, one participant had only worked with personal
repositories, and one participant had worked with personal, academic,
and school repositories.
\dmg{you made it confusing: just say: how many had experience with
  professional repositories and remove the above
}
\evan{Do you want more information about the participants?}

All participants have had at least some experience with the idea of
version control, branching in repositories, and git. The participants
are all from the same lab, and each participant worked with both tools
in the study, thus, keeping the sample populations identical for both
tools, with some variation between participants in experience with
repositories.

\dmg{BIG QUESTION}
for each task, did they do always git tools then linvis for the same commit?  if so, there is potential for learning
bias: they learn the structure of the commits with git tools, so therefore it was easier for them to do
them with linvis.
\dmg{END QUESTION}

\dmg{I'll stop here today}

@@

To mitigate the order bias cased by using one tool before the other or
one commit before the other, we randomize both the order that the tools
are used, and the order that the commits are used, between participants.
The conceptual tasks are run twice, once with \comA and once with \comB,
in the same order for all participants using the git tools.

\evan{This paragraph is kind of tricky to explain well. See if you can
  understand what I'm trying to say} The summarization tasks are run
four times, once with \comA and \tool, \comB and \tool, \comA and git
tools, and \comB and git tools. The participant will start with one
tool, and complete the tasks for both commits in that tool. The order of
the commits is the same as in the conceptual tasks. Once they complete
the summarization tasks for both \comA and \comB, they will switch to
the other tool and complete the tasks again on the commits in the same
order. The order of the tasks is consistent over all four runs. The
questions in the user opinions section do not directly involve the tools
or commits.

>>>>

We conducted a study to quantitatively and qualitatively evaluate the
effectiveness of \tool compared to the DAG-based visualizations found in
Gitk or the git command line. The study is broken into three task-based
sections; conceptual tasks, summarization tasks, and participant
opinions. The study was performed during June of 2017 in a controlled
environment running Ubuntu 14.04. To evaluate the DAG-based
visualizations, participants were allowed to use both Gitk and the git
command-line tools together, which we will refer to as Gitk. \tool was
used for evaluating \mt-based visualizations.

The primary goal of the study is two fold; first determining if the
DAG-based visualization is sufficient for conceptual understanding;
second comparing \tool and Gitk to determine which is more capable of
providing users with a summarization of various metrics involved with
integrating a commit into the repository. The conceptual questions are
to determine if the DAG-based visualizations (Gitk and git command line)
provide enough information to determine how a commit, and any related
commits, are merged into a project. \tool displays this information
directly in the form of a tree, so we will not be performing a
comparison on this section, only evaluating the results from Gitk. The
summarization tasks compare \tool and Gitk, with the participants using
one tool followed by the other on two commits from \mt{s} of
differing sizes. In both of these sections, the participant is given a
task that they are to find an answer to; we evaluate the time it took,
if the answer was correct, and how far from the right answer they were.
The user opinions provide us with a clear comparison between the tools
from the point-of-view of the participant, providing us information
about which aspects of each tool they preferred.

@@

@@

\textbf{The conceptual tasks} are designed to gain better insight on the
comprehension of the visualization in Gitk, and determine if it is
sufficient for understanding what other commits are related to a given
commit, and how those commits are merged. The tasks are outlined in
Table~\ref{tab:conceptual_tasks}. The tasks are performed in order using
Gitk. T1 has the participant draw a diagram of the merge tree,
which will be used to answer the other two questions. This enables us to
see issues with comprehending the DAG visualization. Participants are
given 10 minutes to perform this task, which also is used to familiarize
the participant with the Gitk interface, if they are not already, and
with the set of commits they will be working with. T2 and T3 are drawn
directly from the diagram from T1, giving quantitative metrics to
measure the comprehension. T2 asks the participant to verify the number
of commits, and T3 asks for the number of merges. The three tasks are
performed on both commits before continuing to the summarization tasks.

\textbf{The summarization tasks}, listed in
Table~\ref{tab:summarization_tasks}, are designed to compare the ability
of the participants to summarize information about the two commits used
in the study when using Gitk versus \tool. The tasks are broken into
four sets based on what is being summarized: merges, authorship, files,
and modules. The order that the participants perform each task is
randomized within each task set, and the order of the task sets are then
randomized with the exception of the merge task set which will always
come first. The order that the tools are used is kept consistent for a
participant, but randomized between participants. We measure and
evaluate three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

While the merge task set is part of the summarization task group, it
focuses less with summarization of the commits and more with
comprehension of the model visualizations. The task is designed with two
goals in mind; first we want to measure and define which commits and
which merges the participant has defined to be in the merge tree; second
to help solidify which commits and merges will be used in the merge
tree.

T4 provides a concrete series of merges that the participant believes to
be integrating the commit into the repository. T5 provides a concrete
set of commits that are related to the commit. We place these tasks
first as the answers to these will be necessary for the participant to
answer the tasks that follow, specifically finding the merge into the
master branch. T6 and T7 are related to the authorship of a merge,
determining how many authors are involved and who made the most changes.
T8 and T9 are related to the files that were modified in the merge,
determining how many files were modified and which file had the most
changes. T10 is related to the modules that were modified in the merge.
Modules are not explicitly defined in git, but are a property of the
Linux repository that we noticed while looking at various commits. To
find the module name, take the text up to the first colon in the git log
preview. For the log preview \textit{``ALSA: kernel docs: fix
  sound/core/ kernel-doc''} the module is \textit{``ALSA''}, e.g.

The correctness metric is simply whether the answer was correct or not;
an accuracy of zero indicates a correct answer. The accuracy metrics are
measured slightly differently between tasks due to different
requirements and structures in the answer. In task T4, both the order
and the merges are important. We define the distance to be the edit
distance with insertion, deletion, and transposition at unit cost. The
accuracy is the minimum edit distance necessary to make the answer
correct. T5, T9, and T10 are concerned with the set of elements in the
response; the order is not important. We use the Levenshtein distance,
with insertion, deletion, and substitution operations of unit cost, from
the provided answer to the correct answer. The answers to T6 and T8 are
single number values; the distance is measured as the absolute value of
the difference between the provided answer and the correct answer. T7
has only a single correct answer, thus the distance would either be zero
or one. For this reason, it is acceptable to omit the accuracy metric
from this task.

@@

We include two commits in the study to detect differences between \mt
sizes. The order that the commits are presented to each participant is
randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
non-root nodes. 75\% of the trees contain at least 51 nodes, and the
largest tree contains 7217 nodes.

From this information, we selected one tree with a single commit, and
one tree with seven non-root nodes to represent a small tree and a
medium-sized tree. A majority of the trees are flat, where all nodes
merge directly into the root node. Of the 8031 trees containing at least
seven non-root nodes, only 593 contained at least a single internal
merge node. Trivially, trees with a single node cannot have any internal
nodes, and of the 624 trees with seven non-root nodes, only 135
contained at least one inner node. To increase the complexity of the
medium-sized tree, we randomly selected one of the 135 trees to
represent the medium tree.

From the 2008 trees containing only a single node, we selected one at
random, and trivially selected the only commit in the tree. From the
trees containing seven non-root nodes, we selected one tree randomly,
then randomly selected one of the commits in the tree. We selected
commit \emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree
containing a single node, which we will refer to as \comA, and commit
\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
seven nodes, which we will refer to as \comB. A visual representation
for these trees is shown in Figure~\ref{fig:study_commits}.

\begin{figure}[bpt]
  \centering
  \begin{tabular}{ m{1.5cm} m{3cm} }
    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
  \end{tabular}
  \caption{The merge trees used in the user study,
    containing one and seven non-root nodes, respectively.}
  \label{fig:study_commits}
\end{figure}

@@

The study was conducted with 12 participants, all of whom were masters,
PhD, or post-doc researchers in the field of software engineering. The
participants had between 6 months and 10 years experience with git, with
the median being 3.5 years. Most participants had additional experience
with SVN and CVS.\@ One of the participants in the study worked as a
release engineer, studying merge practices to determine the best way to
merge branches while minimizing the number of merge conflicts, on SVN
repositories. The participants worked with repositories ranging from
around 10 commits up to 38000 commits, with the median being 350
commits. Two of the participants had never collaborated with anyone in a
repository, while the rest had some experience with repositories being
modified by multiple people, with the most being 219. The median number
of collaborators was four. Participants had most experience with
personal and academic repositories. Three of the twelve participants had
experience with professional repositories.

All participants have had at least some experience with version control,
branching in repositories, and git. The participants are from the same
lab, and each participant worked with both tools in the study, thus,
keeping the sample populations identical for both tools, with some
variation between participants in experience with repositories.

@@

To mitigate the order bias cased by using one tool before the other or
one commit before the other, we randomize both the order that the tools
are used, and the order that the commits are used, between participants.
The conceptual tasks are run twice, once with \comA and once with \comB,
in the same order for all participants using the Gitk.

\evan{This paragraph is kind of tricky to explain well. See if you can
  understand what I'm trying to say} The summarization tasks are run
four times, once with \comA and \tool, \comB and \tool, \comA and git
tools, and \comB and Gitk. The participant will start with one
tool, and complete the tasks for both commits in that tool. The order of
the commits is the same as in the conceptual tasks. Once they complete
the summarization tasks for both \comA and \comB, they will switch to
the other tool and complete the tasks again on the commits in the same
order. The order of the tasks is consistent over all four runs. The
questions in the user opinions section do not directly involve the tools
or commits.
